Supporting Information to: Species sympatry shapes brain size evolution in Primates

### 2022-01-28 Article/Supplementary_material_forsubmission_wordworking.Rmd
Update for submission - Separation Sup mat and Main text - Add Rmd word compilation

output: 
 #      - '--lua-filter=scholarly-metadata.lua'
 #      - '--lua-filter=author-info-blocks.lua'
 bookdown::pdf_book:
 #bookdown::word_document2:
   number_sections: false
   toc: false
   citation_package: biblatex
   latex_engine: pdflatex
   fig_caption: true
   #pandoc_args:
     #- '--lua-filter=scholarly-metadata.lua'
     #- '--lua-filter=author-info-blocks.lua'
 keep_tex: true
always_allow_html: true

@@

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(include=FALSE, echo=FALSE, message=FALSE)
knitr::opts_chunk$set(dpi=300) # Fig. resolution and size out.width = '100%', 
knitr::opts_chunk$set(fig.pos = 'p'#, fig.align = 'center'
                      ) # Places figures on pages separate from text, centered
knitr::opts_chunk$set(fig.env="figure") # Latex Fig. environment
knitr::opts_knit$set(eval.after = "fig.cap") #To insert R code into R Fig. caption
```

@@

#Plot
library(RColorBrewer)
library(tidyr)
library(stringr)
library(svMisc)
library(plotrix)
library(circlize)

@@

library(tidyverse)
#Import own function
source("T:/Saved_PhD/Empirical_analysis/Scripts&Functions/Functions/toolbox.R", local = knitr::knit_global())

@@

par(mar=c(3, 5, 2, 1), mgp=c(2, 0.5, 0), xpd=TRUE)
plot(
  x=0, y=0, xlab="Overlap threshold", ylab="Percentage of species with different biogeographic areas\ncompared to when a 15% overlap is used", 
  xlim=c(thresholdPresenceRange[1],thresholdPresenceRange[length(thresholdPresenceRange)]), ylim=c(0,0.4),
  las=1, type="n", tcl=-0.25, frame.plot=FALSE, 
  xaxt="n",xaxs="i",yaxs="i", yaxt="n")

@@

|   Second, for each species, trait estimates could vary slightly among datasets (see Fig. Appendix Fig. S2). Particularly, although correlations between measures from the different datasets seem good enough, it existed a variation in absolute measurements (Fig. Appendix Fig. S2). We are aware that this could contribute to blurring results [@Wartel2019; @Hooper2021]. In addition, this study is based on several arbitrary thresholds, namely (i) to assess species sympatry (see Appendix Fig. S1) and (ii) to assess the species dietary guild (see Fig. Appendix Fig. S2) which can cause sensitivity of the results to the chosen parameters. To account for these three sources of variability, we refitted several times the six models of trait evolution (BM, OU, EB, MC, DD$_{lin}$ and DD$_{exp}$) with (1) random samples of the dietary and brain traits in case of multiple values available (i.e. equal probability for each possible value to be selected), (2) used the low or high threshold for assessing frugivory, folivory, and species sympatry, and (3) various biogeography and dietary evolutionary history reconstructions.

@@

```{=latex}
\begin{figure}
\centering\includegraphics[width=0.65\linewidth]{C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/selectionGradientPGLS.pdf}
\caption{\scriptsize{Phylogenetic regressions of relative brain size as a function of sympatry level indices | Left graphics show the effect of the number of sympatric species on the brain size, when the effect of the average percentage of overlapping range with sympatric frugivorous species  is averaged, while the right graphics do the opposite. Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\% confidence interval is depicted by the transparent grey background.}}\label{fig:figRegressionGradient}
\end{figure}
```
\clearpage

@@

```{=latex}
\begin{figure}
\centering\includegraphics[width=0.7\linewidth]{C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationPGLS.pdf}
\caption{\scriptsize{Phylogenetic regressions of the net diversification rate as a function of the size of the different brain areas | Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\% highest density posterior is depicted by the transparent grey background.}}\label{fig:figRegressionDiversification}
\end{figure}
```

@@

```{=latex}
\begin{figure}
\centering\includegraphics[width=0.65\linewidth]{C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationAndSympatryPGLS.pdf}
\caption{\scriptsize{Phylogenetic regressions of the net diversification rate as a function of sympatry level indices | The left graphic depicts the effect of the number of sympatric species on the brain size, when the effect of the average percentage of overlapping range with sympatric frugivorous species is averaged, while the right graphic does the opposite. Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\% confidence interval is depicted by the transparent grey background.}}\label{fig:figRegressionDiversificationSympatry}
\end{figure}
```

@@

rownames(summarySensitivityGradient) <- NULL
knitr::kable(summarySensitivityGradient, escape=TRUE, booktabs = TRUE, #Remove brain 1:3
             caption = "Sensitivity analysis of phylogenetic regressions to assess the relationship between relative brain sizes and species sympatry | Depicted is the minimum and maximum of estimates when one observation was removed at a time (DfBetas) or when varying the used phylogenetic tree and the data sampling (Phylogeny/Data).") %>% 
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::kable_styling(latex_options="scale_down") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position") %>%
  kableExtra::add_header_above(c("Regression" = 2, "DfBetas" = 3, "Phylogeny/Data" = 3))
```

@@

knitr::kable(summarySensitivityDiversification[,-c(3:5)], escape=TRUE, booktabs = TRUE, #Remove brain 1:3
             caption = "Sensitivity analysis of phylogenetic regressions to assess the relationship between species diversification and relative brain sizes | Depicted is the minimum and maximum of estimates when varying the used phylogenetic tree and the data sampling (Phylogeny/Data) or when the sampling fraction varied (Sampling fraction).") %>% 
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::kable_styling(latex_options="scale_down") %>%
  kableExtra::kable_styling(latex_options = "HOLD_position") %>%
  kableExtra::add_header_above(c("Regression" = 2, "Phylogeny/Data" = 3, "Sampling fraction" = 3))
```

@@

```{r tabledfsensitivity3, include=TRUE}
summarySensitivitydiversificationAndSympatry[,1] <- gsub("Overlap", "\\% of overlapping range", summarySensitivitydiversificationAndSympatry[,1])
summarySensitivitydiversificationAndSympatry[,1] <- gsub("N co-occurrence", "Number of sympatric frugivores", summarySensitivitydiversificationAndSympatry[,1])

knitr::kable(summarySensitivitydiversificationAndSympatry, escape=TRUE, booktabs = TRUE,
             caption = "Sensitivity analysis of phylogenetic regressions to assess the relationship between species diversification and sympatry | Depicted is the minimum and maximum of estimates when one observation was removed at a time (DfBetas), when varying the used phylogenetic tree and the data sampling (Phylogeny/Data), or when the sampling fraction varied (Sampling fraction)") %>% 
  kableExtra::kable_styling(latex_options = "striped") %>%
  kableExtra::kable_styling(latex_options="scale_down") %>%
  kableExtra::add_header_above(c("Model" = 1, "DfBetas" = 3, "Phylogeny/Data" = 3, "Sampling fraction" = 3))

>>>>

output: 
 #      - '--lua-filter=scholarly-metadata.lua'
 #      - '--lua-filter=author-info-blocks.lua'
 #bookdown::pdf_book:
 bookdown::word_document2:
   number_sections: false
   toc: false
   citation_package: biblatex
   latex_engine: pdflatex
   fig_caption: true
   #pandoc_args:
     #- '--lua-filter=scholarly-metadata.lua'
     #- '--lua-filter=author-info-blocks.lua'
 keep_tex: true
always_allow_html: true

@@

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(include=FALSE, echo=FALSE, message=FALSE)
knitr::opts_chunk$set(dpi=300) # Figure resolution and size out.width = '100%', 
knitr::opts_chunk$set(fig.pos = 'p'#, fig.align = 'center'
                      ) # Places figures on pages separate from text, centered
knitr::opts_chunk$set(fig.env="figure") # Latex figure environment
knitr::opts_knit$set(eval.after = "fig.cap") #To insert R code into R figure caption
```

<!-- TC:ignore -->

@@

#Plot
library(RColorBrewer)
library(tidyr)
library(stringr)
library(svMisc)
library(plotrix)
library(circlize)
library(magick)
library(knitr)
library(png)

@@

#Data handling
library(tidyverse)
library(huxtable)

#Import own function
source("T:/Saved_PhD/Empirical_analysis/Scripts&Functions/Functions/toolbox.R", local = knitr::knit_global())

@@

par(mar=c(3, 5, 2, 1), mgp=c(2, 0.5, 0), xpd=TRUE)
plot(
  x=0, y=0, xlab="Overlap threshold", ylab="Percentage of species with different biogeographic\nareas compared to when a 15% overlap is used", 
  xlim=c(thresholdPresenceRange[1],thresholdPresenceRange[length(thresholdPresenceRange)]), ylim=c(0,0.4),
  las=1, type="n", tcl=-0.25, frame.plot=FALSE, 
  xaxt="n",xaxs="i",yaxs="i", yaxt="n")

@@

|   Second, for each species, trait estimates could vary slightly among datasets (see Appendix Fig. S2). Particularly, although correlations between measures from the different datasets seem good enough, it existed a variation in absolute measurements (Appendix Fig. S2). We are aware that this could contribute to blurring results [@Wartel2019; @Hooper2021]. In addition, this study is based on several arbitrary thresholds, namely (i) to assess species sympatry (see Appendix Fig. S1) and (ii) to assess the species dietary guild (see Appendix Fig. S2) which can cause sensitivity of the results to the chosen parameters. To account for these three sources of variability, we refitted several times the six models of trait evolution (BM, OU, EB, MC, DD$_{lin}$ and DD$_{exp}$) with (1) random samples of the dietary and brain traits in case of multiple values available (i.e. equal probability for each possible value to be selected), (2) used the low or high threshold for assessing frugivory, folivory, and species sympatry, and (3) various biogeography and dietary evolutionary history reconstructions.

@@

<!-- ```{=latex} -->
<!-- \begin{figure} -->
<!-- \centering\includegraphics[width=0.65\linewidth]{C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/selectionGradientPGLS.pdf} -->
<!-- \caption{\scriptsize{Phylogenetic regressions of relative brain size as a function of sympatry level indices | Left graphics show the effect of the number of sympatric species on the brain size, when the effect of the average percentage of overlapping range with sympatric frugivorous species  is averaged, while the right graphics do the opposite. Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\% confidence interval is depicted by the transparent grey background.}}\label{fig:figRegressionGradient} -->
<!-- \end{figure} -->
<!-- ``` -->

```{r}
error_file = magick::image_read("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/selectionGradientPGLS.pdf", density=500)
error_file = magick::image_convert(error_file, "png")
image_write(error_file, path = "C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/selectionGradientPGLS.png", format = "png")
```

```{r figRegressionGradient, include=TRUE, fig.width = 7, fig.height = 7, fig.cap= "Phylogenetic regressions of relative brain size as a function of sympatry level indices | Left graphics show the effect of the number of sympatric species on the brain size, when the effect of the average percentage of overlapping range with sympatric frugivorous species  is averaged, while the right graphics do the opposite. Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\\% confidence interval is depicted by the transparent grey background."}

knitr::include_graphics("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/selectionGradientPGLS.png", error=TRUE)

@@

<!-- ```{=latex} -->
<!-- \begin{figure} -->
<!-- \centering\includegraphics[width=0.7\linewidth]{C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationPGLS.pdf} -->
<!-- \caption{\scriptsize{Phylogenetic regressions of the net diversification rate as a function of the size of the different brain areas | Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\% highest density posterior is depicted by the transparent grey background.}}\label{fig:figRegressionDiversification} -->
<!-- \end{figure} -->
<!-- ``` -->

```{r}
error_file = magick::image_read("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationPGLS.pdf", density=500)
error_file = magick::image_convert(error_file, "png")
image_write(error_file, path = "C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationPGLS.png", format = "png")
```

```{r figRegressionDiversification, include=TRUE, fig.width = 7, fig.height = 7, fig.cap= "Phylogenetic regressions of the net diversification rate as a function of the size of the different brain areas | Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\\% highest density posterior is depicted by the transparent grey background."}

knitr::include_graphics("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationPGLS.png", error=TRUE)

@@

<!-- ```{=latex} -->
<!-- \begin{figure} -->
<!-- \centering\includegraphics[width=0.65\linewidth]{C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationAndSympatryPGLS.pdf} -->
<!-- \caption{\scriptsize{Phylogenetic regressions of the net diversification rate as a function of sympatry level indices | The left graphic depicts the effect of the number of sympatric species on the brain size, when the effect of the average percentage of overlapping range with sympatric frugivorous species is averaged, while the right graphic does the opposite. Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\% confidence interval is depicted by the transparent grey background.}}\label{fig:figRegressionDiversificationSympatry} -->
<!-- \end{figure} -->
<!-- ``` -->

```{r}
error_file = magick::image_read("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationAndSympatryPGLS.pdf", density=500)
error_file = magick::image_convert(error_file, "png")
image_write(error_file, path = "C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationAndSympatryPGLS.png", format = "png")
```

```{r figRegressionDiversificationSympatry, include=TRUE, fig.width = 7, fig.height = 7, fig.cap= "Phylogenetic regressions of the net diversification rate as a function of sympatry level indices | The left graphic depicts the effect of the number of sympatric species on the brain size, when the effect of the average percentage of overlapping range with sympatric frugivorous species is averaged, while the right graphic does the opposite. Raw data are depicted with points, while the segments that link them correspond to the projected phylogenetic tree. The model fit is shown with the plain black line and the associated 95\\% confidence interval is depicted by the transparent grey background."}

knitr::include_graphics("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Plots/diversificationAndSympatryPGLS.png", error=TRUE)

@@

# knitr::kable(summarySensitivityGradient, escape=TRUE, booktabs = TRUE, #Remove brain 1:3
#              caption = "Sensitivity analysis of phylogenetic regressions to assess the relationship between relative brain sizes and species sympatry | Depicted is the minimum and maximum of estimates when one observation was removed at a time (DfBetas) or when varying the used phylogenetic tree and the data sampling (Phylogeny/Data).") %>% 
#   kableExtra::kable_styling(latex_options = "striped") %>%
#   kableExtra::kable_styling(latex_options="scale_down") %>%
#   kableExtra::kable_styling(latex_options = "HOLD_position") %>%
#   kableExtra::add_header_above(c("Regression" = 2, "DfBetas" = 3, "Phylogeny/Data" = 3))

summarySensitivityGradient.toPlot <- rbind(colnames(summarySensitivityGradient), summarySensitivityGradient)
colnames(summarySensitivityGradient.toPlot) <- 1:ncol(summarySensitivityGradient.toPlot)
rownames(summarySensitivityGradient) <- NULL

as_huxtable(summarySensitivityGradient.toPlot, add_colnames = FALSE, add_rownames = FALSE) %>%
  set_caption("Sensitivity analysis of phylogenetic regressions to assess the relationship between relative brain sizes and species sympatry | Depicted is the minimum and maximum of estimates when one observation was removed at a time (DfBetas) or when varying the used phylogenetic tree and the data sampling (Phylogeny/Data).") %>%

  #Main layout
  set_bottom_border(row = 1, col = everywhere) %>%
  set_bold(row = 1, col = everywhere) %>%
  set_background_color(evens, everywhere, "grey95") %>%

  #Sub-headers
  insert_row("Regression", "", "DfBetas", "", "", "Phylogeny/Data", "", "", after = 0) %>% 
  merge_cells(1, 1:2) %>% 
  merge_cells(1, 3:5) %>% 
  merge_cells(1, 6:8) %>% 
  set_align(1, everywhere, "center") %>% 
  set_bold(1, everywhere) %>%  
  as_flextable()

@@

# knitr::kable(summarySensitivityDiversification[,-c(3:5)], escape=TRUE, booktabs = TRUE, #Remove brain 1:3
#              caption = "Sensitivity analysis of phylogenetic regressions to assess the relationship between species diversification and relative brain sizes | Depicted is the minimum and maximum of estimates when varying the used phylogenetic tree and the data sampling (Phylogeny/Data) or when the sampling fraction varied (Sampling fraction).") %>% 
#   kableExtra::kable_styling(latex_options = "striped") %>%
#   kableExtra::kable_styling(latex_options="scale_down") %>%
#   kableExtra::kable_styling(latex_options = "HOLD_position") %>%
#   kableExtra::add_header_above(c("Regression" = 2, "Phylogeny/Data" = 3, "Sampling fraction" = 3))

summarySensitivityDiversification.toPlot <- rbind(colnames(summarySensitivityDiversification), summarySensitivityDiversification)
colnames(summarySensitivityDiversification.toPlot) <- 1:ncol(summarySensitivityDiversification.toPlot)

as_huxtable(summarySensitivityDiversification.toPlot[,-c(3:5)], add_colnames = FALSE, add_rownames = FALSE) %>%
  set_caption( "Sensitivity analysis of phylogenetic regressions to assess the relationship between species diversification and relative brain sizes | Depicted is the minimum and maximum of estimates when varying the used phylogenetic tree and the data sampling (Phylogeny/Data) or when the sampling fraction varied (Sampling fraction).") %>%

  #Main layout
  set_bottom_border(row = 1, col = everywhere) %>%
  set_bold(row = 1, col = everywhere) %>%
  set_background_color(evens, everywhere, "grey95") %>%

  #Sub-headers
  insert_row("Regression", "","Phylogeny/Data", "", "", "Sampling fraction", "", "", after = 0) %>% 
  merge_cells(1, 1:2) %>% 
  merge_cells(1, 3:5) %>% 
  merge_cells(1, 6:8) %>% 
  set_align(1, everywhere, "center") %>% 
  set_bold(1, everywhere) %>%  
  as_flextable()
```

@@

```{r tabledfsensitivity3, include=TRUE}
summarySensitivitydiversificationAndSympatry[,1] <- gsub("Overlap", "\\% of overlapping range", summarySensitivitydiversificationAndSympatry[,1])
summarySensitivitydiversificationAndSympatry[,1] <- gsub("N co-occurrence", "Number of sympatric frugivores", summarySensitivitydiversificationAndSympatry[,1])
# 
# knitr::kable(summarySensitivitydiversificationAndSympatry, escape=TRUE, booktabs = TRUE,
#              caption = "Sensitivity analysis of phylogenetic regressions to assess the relationship between species diversification and sympatry | Depicted is the minimum and maximum of estimates when one observation was removed at a time (DfBetas), when varying the used phylogenetic tree and the data sampling (Phylogeny/Data), or when the sampling fraction varied (Sampling fraction)") %>% 
#   kableExtra::kable_styling(latex_options = "striped") %>%
#   kableExtra::kable_styling(latex_options="scale_down") %>%
#   kableExtra::add_header_above(c("Model" = 1, "DfBetas" = 3, "Phylogeny/Data" = 3, "Sampling fraction" = 3))

summarySensitivitydiversificationAndSympatry.toPlot <- rbind(colnames(summarySensitivitydiversificationAndSympatry), summarySensitivitydiversificationAndSympatry)
colnames(summarySensitivitydiversificationAndSympatry.toPlot) <- 1:ncol(summarySensitivitydiversificationAndSympatry.toPlot)

as_huxtable(summarySensitivitydiversificationAndSympatry.toPlot, add_colnames = FALSE, add_rownames = FALSE) %>%
  set_caption(  "Sensitivity analysis of phylogenetic regressions to assess the relationship between species diversification and sympatry | Depicted is the minimum and maximum of estimates when one observation was removed at a time (DfBetas), when varying the used phylogenetic tree and the data sampling (Phylogeny/Data), or when the sampling fraction varied (Sampling fraction)") %>%

  #Main layout
  set_bottom_border(row = 1, col = everywhere) %>%
  set_bold(row = 1, col = everywhere) %>%
  set_background_color(evens, everywhere, "grey95") %>%

  #Sub-headers
  insert_row("Regression", "DfBetas", "", "","Phylogeny/Data", "", "", "Sampling fraction", "", "", after = 0) %>%
  merge_cells(1, 2:4) %>% 
  merge_cells(1, 5:7) %>% 
  merge_cells(1, 8:10) %>% 
  set_align(1, everywhere, "center") %>% 
  set_bold(1, everywhere) %>%  
  as_flextable()
